\documentclass[11pt]{article}

%TC:ignore
\input{config.tex}


\begin{document}

% Roman page numbers for preamble
\pagenumbering{roman}

% Title page
\input{contents/title/titlepage.tex}


% Executive summary
\section*{Executive summary}
This project report presents the design, development and deployment of a
complete IoT weather station network for use in agricultural settings. The
network was built from the ground up with off-the-shelf electronics and a custom
weather proofed enclosure. The system streams one minute sensor measurements to
a linked web application also developed in this project called Agriscanner
(\url{https://agriscanner.onrender.com}).

The Agriscanner webapp provides users with historical, real-time, and forecast
weather data through an intuitive and accessible interface. The webapp uses a
machine learning model (LightGBM), to display microclimate specific forecast
data that is trained by comparing general forecast information to sensor
readings from the IoT network.

General forecasting models often miss smaller climate variations that exist in
farms and fields. While commercial weather monitoring stations are often a
trade-off between the range they can transmit information and their price.
Because farms are remote, any device collecting information must be
self-sustaining in power and able to send data over long distances.

The weather stations in this design achieved a measured range of 1,200m from the
receiver in non-ideal circumstances and the final design is estimated to have
over double the range of commercial options due to the inclusion of a repeater.

On the software side, the backend was built with a relational database that
keeps a log of all sensor readings, along with a secure API that facilitates
these database queries. It serves a front-end web interface that scored highly
for usability on both mobile and desktop.

The machine learning model's forecast was compared to actual sensor data and
showed that it was more accurate than the general forecast in most scenarios,
but more training data is needed to improve its accuracy further.

It was an ambitious project for the timescale - involving the design and
construction of four hardware devices; the development of a backend database and
API; the development of frontend web interface and the training and deployment
of a set of machine learning models.

\addcontentsline{toc}{section}{Executive summary}


% Dedication and acknowledgments
\newpage
\section*{Dedication and acknowledgments}
I would like to thank my parents for the use of their garden, their wisdom and
their near infinite patience; my supervisor for her advice and support
throughout this project; and my partner, for helping me with range testing the
network by walking the length of the Downs with a box of electronics, despite
strange looks from passers-by.


\addcontentsline{toc}{section}{Dedication and acknowledgments}


% Declaration
\newpage
\section*{Author's declaration}
I declare that the work in this dissertation was carried out in accordance with
the requirements of the University's Regulations and Code of Practice for
Research Degree Programmes and that it has not been submitted for any other
academic award. Except where indicated by specific reference in the text, the
work is the candidate's own work. Work done in collaboration with, or with the
assistance of, others, is indicated as such. Any views expressed in the
dissertation are those of the author.


SIGNED:\qquad Adam Sidnell \qquad \qquad \qquad DATE:\qquad    2nd September
2025
\addcontentsline{toc}{section}{Author's declaration}


% Contents
\clearpage
\begingroup
\pagestyle{plain}
\vspace*{3cm}
\tableofcontents
\clearpage
\endgroup

% Figures
\clearpage
\begingroup
\pagestyle{plain}
\vspace*{3cm}
\listoffigures
\clearpage
\endgroup
%TC:endignore

% Use arabic numbers for main contents pages
\pagenumbering{arabic}

% Introduction
\input{contents/part-1/introduction.tex}

% Background
\part{Background}

% Internet of things

\input{contents/part-1/lora.tex}

\input{contents/part-1/iot.tex}

% Microclimates
\input{contents/part-1/microclimates.tex}

\part{Hardware development}

\input{contents/part-2/hw-overview.tex}

\input{contents/part-2/hw-design.tex}

\input{contents/part-2/hw-development.tex}

\input{contents/part-2/hw-deployment.tex}

\part{Software development}

\input{contents/part-3/hardware-programming.tex}

\input{contents/part-3/sw-overview.tex}

\input{contents/part-3/website.tex}

\input{contents/part-3/machine-learning.tex}

\part{Evaluation}

\input{contents/part-4/commercial-comparison.tex}

\input{contents/part-4/web-app.tex}

\input{contents/part-4/machine-learning.tex}

\input{contents/part-4/conclusion.tex}


% References
\newpage
\addcontentsline{toc}{section}{References}
\printbibliography

\appendix
\newpage
\part{Appendices}

% Override spacing just for appendix
\makeatletter
\patchcmd{\section}{\clearpage\vspace*{3cm}\thispagestyle{plain}}{\clearpage\thispagestyle{plain}}{}{}
\titleformat{\part}[block]
{\normalfont\centering\bfseries} {\LARGE \partname\ \thepart}{1em}{\Huge}
[\thispagestyle{plain}\clearpage]
\makeatother

%TC:ignore
\input{contents/appendix/appendix.tex}
%TC:endignore

\end{document}
